---
title: "Corrected TEs Analysis"
author: "Nel Marín"
date: "20/6/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Import mapping files

```{r Files, echo=FALSE}

#Set the working directory
setwd("~/TFM/12_Corrected_results")

#Load the TSV file using read.delim()
dataframe_complete <- read.delim("~/TFM/5_TEs_mapping_&_repeat_analysis/thedataframe_complete.tsv", header = FALSE, stringsAsFactors = FALSE)

#Change the column names
colnames(dataframe_complete) <- c("seqname", "source", "feature", "start", "end",
                                  "score", "strand", "frame", "class", "family",
                                  "group", "H1.0", "H1.2",  "H1.3", "H1.4", "H1.5",
                                  "H1X","H3K9me3", "H3K27me3",
                                  "K9me3_AJV7", "K9me3_AJV84X")

```

## 2. Create the actual dataframe

```{r Dataframe}

#Remove the H3K9me3 columns that we are not using
new_dataframe_complete <- dataframe_complete[, c(1:19)]

#New name for the H3K9me3 sample
#colnames(new_dataframe_complete)[18] <- "K9me3"

#Identify numeric columns
numeric_cols <- sapply(new_dataframe_complete, is.numeric)

#Change "start", "end" & "score" to FALSE (not numeric)
numeric_cols[4:6] <- FALSE

#Filter the dataframe
new_filtered_dataframe_complete <- new_dataframe_complete[rowSums(new_dataframe_complete[, numeric_cols] != 0) > 0, ]
rownames(new_filtered_dataframe_complete) <- seq(1:nrow(new_filtered_dataframe_complete))

```

## 3. Calculate the median abundance

In this step, we will calculate the median abundance of each sample at the different levels of organization (class, family & group). In order to achieve this, we will create three dataframes, one for each level, where the rows will be the different classes, families or groups and the columns will be ten samples (histones & histone marks).

#### 3.1. Class

```{r Class abundance}

#Create an empty dataframe to store the results
df_class <- data.frame(matrix(nrow = 10, ncol = 8))
#Set the columns names
colnames(df_class) <- c("H1.0", "H1.2",  "H1.3", "H1.4", "H1.5", "H1X",
                        "H3K9me3", "H3K27me3")
#Extract the repeat classes and set as row names
repeat_classes <- sort(unique(new_filtered_dataframe_complete$class))
rownames(df_class) <- repeat_classes

#Loop to calculate median values of each class and sample
for (i in 1:length(repeat_classes)) {
  current_class <- repeat_classes[i]
  class_subset <- new_filtered_dataframe_complete[new_filtered_dataframe_complete$class == current_class, 12:19]
  median_values <- apply(class_subset, 2, median)
  df_class[i, ] <- c(median_values)
}

#Test if the results are correct
#test_LINE <- dataframe_complete[dataframe_complete$class == "LINE", ]
#summary(test_LINE)

```

#### 3.2. Family

```{r Family abundance}

#Create an empty dataframe to store the results
df_family <- data.frame(matrix(nrow = 50, ncol = 8))
#Set the columns names
colnames(df_family) <- c("H1.0", "H1.2",  "H1.3", "H1.4", "H1.5", "H1X",
                         "H3K9me3", "H3K27me3")
#Extract the repeat families and set as row names
repeat_families <- sort(unique(new_filtered_dataframe_complete$family))
rownames(df_family) <- repeat_families

#Loop to calculate median values of each family and sample
for (i in 1:length(repeat_families)) {
  current_family <- repeat_families[i]
  family_subset <- new_filtered_dataframe_complete[new_filtered_dataframe_complete$family == current_family, 12:19]
  median_values <- apply(family_subset, 2, median)
  df_family[i, ] <- c(median_values)
}

#Test if the results are correct
#test_Alu <- dataframe_complete[dataframe_complete$family == "Alu", ]
#summary(test_Alu)

```

#### 3.3. Group

```{r Group abundance}

#Extract the repeat families and set as row names
repeat_groups <- sort(unique(new_filtered_dataframe_complete$group))
rownames(df_group) <- repeat_groups

#Create an empty dataframe to store the results
df_group <- data.frame(matrix(nrow = length(repeat_groups), ncol = 8))
#Set the columns names
colnames(df_group) <- c("H1.0", "H1.2",  "H1.3", "H1.4", "H1.5", "H1X",
                        "H3K9me3", "H3K27me3")

#Loop to calculate median values of each family and sample
for (i in 1:length(repeat_groups)) {
  current_group <- repeat_groups[i]
  group_subset <- new_filtered_dataframe_complete[new_filtered_dataframe_complete$group == current_group, 12:19]
  median_values <- apply(group_subset, 2, median)
  df_group[i, ] <- c(median_values)
}

#Test if the results are correct
#test_AluJo <- dataframe_complete[dataframe_complete$group == "AluJo", ]
#summary(test_AluJo)

```

## 4. Scale the columns

Before creating the heatmaps, it is necessary to scale the columns of the new dataframes that contain numerical information (all of them).

```{r Scaling}

#Scale the new dataframes
df_class_scaled <- scale(df_class)
df_family_scaled <- scale(df_family)
df_group_scaled <- scale(df_group)

```

## 5. Create the heatmaps

#### 5.1. Class heatmap \[DONE\]

```{r Class heatmap}

library(ggplot2)
library(pheatmap)

#Create the annotation
annotation_class <- data.frame(Class = repeat_classes)
rownames(annotation_class) <- rownames(df_class_scaled)

#Make sure the levels of the Class variable match the colors
annotation_class$Class <- factor(annotation_class$Class, 
                                 levels = unique(annotation_class$Class))

#Define colors for each class
class_colors <- c("#FB83DD", "#4DB8FF", "#C8C0A8", "#9898FC", "#C8F63F",
                  "#6C99DA", "#FF4444", "#5EF49E", "#FFD700", "#FE934B")
#class_colors <- rainbow(10)
#class_colors[3] <- "#FFF700"
#class_colors[5] <- "#00A642"
#class_colors[7] <- "#008CFF"
#class_colors[8] <- "#0400FF"
#class_colors[9] <- "#9900FF"

#Set the annotation colors
annotation_class_colors <- list(Class = setNames(class_colors, levels(annotation_class$Class)))

#Plot the heatmap
c <- pheatmap(df_class_scaled, 
              color = colorRampPalette(c("blue3", "white", "firebrick1"))(100),
              cutree_cols = 2, 
              cutree_rows = 4,
              annotation_row = annotation_class,
              annotation_colors = annotation_class_colors)

#Plot the heatmap without the samples K9me3_AJV7 & K9me3_AJV84X
#df_class_scaled_filtered <- df_class_scaled[, -c(9, 10)]
#C <- pheatmap(df_class_scaled_filtered, 
#              color = colorRampPalette(c("blue3", "white", "firebrick1"))(100),
#              cutree_cols = 2, 
#              cutree_rows = 4,
#              annotation_row = annotation_class,
#              annotation_colors = annotation_class_colors)

#Export the plots
ggsave("heatmap_class_FINAL.png", plot = c, width = 10, height = 4, dpi = 300)
#ggsave("heatmap_class_NEW.png", plot = C, width = 10, height = 5, dpi = 300)

```

#### 5.2. Family heatmap \[DONE\]

```{r Family heatmap}

library(ggplot2)
library(pheatmap)

#Create the annotation
class_family <- new_filtered_dataframe_complete[, c(9, 10)]
annotation_family <- subset(class_family, !duplicated(class_family))
annotation_family <- annotation_family[order(annotation_family$class), ]
rownames(annotation_family) <- annotation_family[, 2]
annotation_family <- annotation_family[, -2, drop = FALSE]

#Make sure the levels of the Class variable match the colors
annotation_family$class <- factor(annotation_family$class, 
                                  levels = unique(annotation_family$class))

#Set the annotation colors
annotation_family_colors <- list(
  class = setNames(class_colors, levels(annotation_family$class)))

#Plot the heatmap
f <- pheatmap(df_family_scaled, 
              color = colorRampPalette(c("blue3", "white", "firebrick1"))(100),
              cutree_cols = 2, 
              cutree_rows = 4,
              annotation_row = annotation_family,
              annotation_colors = annotation_family_colors)

#Plot the heatmap without the samples K9me3_AJV7 & K9me3_AJV84X
#df_family_scaled_filtered <- df_family_scaled[, -c(9, 10)]
#P <- pheatmap(df_family_scaled_filtered, 
#              color = colorRampPalette(c("blue3", "white", "firebrick1"))(100),
#              cutree_cols = 2, 
#              cutree_rows = 4,
#              annotation_row = annotation_family,
#              annotation_colors = annotation_family_colors)

#Export the plot
ggsave("heatmap_family.png", plot = f, width = 6, height = 10, dpi = 300)
#ggsave("heatmap_family_NEW.png", plot = P, width = 6, height = 10, dpi = 300)

```

#### 5.3. Group heatmap

##### 5.3.1. Group heatmap split by class

```{r Group heatmap split by class}

library(ggplot2)
library(pheatmap)

#Create the annotation
class_group <- new_filtered_dataframe_complete[, c(9, 11)]
annotation_group <- subset(class_group, !duplicated(class_group))
annotation_group <- annotation_group[order(annotation_group$class), ]
rownames(annotation_group) <- annotation_group[, 2]
annotation_group <- annotation_group[, -2, drop = FALSE]

#Make sure the levels of the Class variable match the colors
annotation_group$class <- factor(annotation_group$class, 
                                  levels = unique(annotation_group$class))

#Set the annotation colors
annotation_group_colors <- list(
  class = setNames(class_colors, levels(annotation_group$class)))

#Prepare the input for the loop
df_merged <- merge(df_group_scaled, annotation_group, 
                   by = "row.names", row.names = 1)
rownames(df_merged) <- df_merged[, 1]
df_merged <- df_merged[, -1, drop = FALSE]
df_list <- split(df_merged, df_merged$class)

#########################################################
#NEW ANNOTATION: FAMILY IN COLOUR GRADIENT
#Use annotation_v2
class_family_group <- new_filtered_dataframe_complete[, c(9, 10, 11)]
annotation_v2 <- subset(class_family_group, !duplicated(class_family_group))
rownames(annotation_v2) <- seq(1,nrow(annotation_v2))

#Family colors
family_colors <- colorRampPalette(c("#FF0000", "white"))(23)
#Check the color gradient
#image(matrix(1:23, ncol=1), col = family_colors)
family_colors <- append(family_colors, colorRampPalette(c("#FF9900", "white"))(9))
family_colors <- append(family_colors, "#FFF700")
family_colors <- append(family_colors, colorRampPalette(c("#33FF00", "white"))(6))
family_colors <- append(family_colors, "#00A642")
family_colors <- append(family_colors, "#00FFFF")
family_colors <- append(family_colors, colorRampPalette(c("#008CFF", "white"))(4))
family_colors <- append(family_colors, colorRampPalette(c("#0400FF", "white"))(3))
family_colors <- append(family_colors, "#9900FF")
family_colors <- append(family_colors, "#FF0099")

#Set the annotation colors
class_family_group <- unique(class_family_group)
class_family_group <- class_family_group[order(class_family_group$class,
                                               class_family_group$family,
                                               class_family_group$group), ]
rownames(class_family_group) <- seq(1, nrow(class_family_group))
class_family_group$family <- factor(class_family_group$family, 
                                    levels = unique(class_family_group$family))
annotation_group_colors_NEW <- list(
  class = setNames(class_colors, levels(annotation_group$class)),
  family = setNames(family_colors, levels(class_family_group$family)))
#########################################################

for (i in 1:length(df_list)) {
  df_class <- df_list[[i]]
  df_class <- df_class[, -ncol(df_class)] #Remove the last column
  displayed_family <- class_family_group[class_family_group$class == names(df_list[i]), "family"]
  displayed_family <- as.vector(unique(displayed_family))
  subannotation_v2 <- annotation_v2[annotation_v2$family %in% displayed_family, ]
  subannotation_group_colors_NEW <- annotation_group_colors_NEW
  subannotation_group_colors_NEW[[2]] <- annotation_group_colors_NEW[[2]][names(annotation_group_colors_NEW[[2]]) %in% displayed_family]
  if (nrow(df_class) == 1) { 
    g <- pheatmap(df_class,
                  main = names(df_list)[i],
                  cluster_rows = FALSE,
                  cutree_cols = 2, 
                  annotation_row = subannotation_v2,
                  annotation_colors = subannotation_group_colors_NEW,
                  color = colorRampPalette(c("blue3", "white", "firebrick1"))(100))
    ggsave(paste0("heatmap_group_C_", names(df_list)[i], ".png"),
           plot = g, dpi = 300, width = 8, height = 3, limitsize = FALSE)
  } else {
    g <- pheatmap(df_class,
                  main = names(df_list)[i],
                  cutree_cols = 2, 
                  annotation_row = subannotation_v2,
                  annotation_colors = subannotation_group_colors_NEW,
                  color = colorRampPalette(c("blue3", "white", "firebrick1"))(100))
    if (nrow(df_class) < 10) {
      ggsave(paste0("heatmap_group_C_", names(df_list)[i], ".png"), 
             plot = g, width = 8, height = 4, dpi = 300)
    } else if (nrow(df_class) < 100) {
      ggsave(paste0("heatmap_group_C_", names(df_list)[i], ".png"), 
             plot = g, width = 16, height = 10, dpi = 300)
    } else {
      ggsave(paste0("heatmap_group_C_", names(df_list)[i], ".png"), 
             plot = g, width = 8, height = 49, dpi = 300)
    }
  }
}
```

##### 4.3.3. Grou
