---
title: "Group_count"
author: "Nel Marín"
date: "17/5/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Import files

Import the files needed from other folder.

```{r Files}

#Set the working directory
setwd("~/TFM/6_Genome_characterization")

#Load the TSV file using read.delim()
dataframe_complete <- read.delim("~/TFM/5_TEs_mapping_&_repeat_analysis/thedataframe_complete.tsv", header = FALSE, stringsAsFactors = FALSE)

#Check the loaded data
#head(dataframe_complete)

#Change the column names
colnames(dataframe_complete) <- c("seqname", "source", "feature", "start", "end",
                                  "score", "strand", "frame", "class", "family",
                                  "group", "H1.0", "H1.2",  "H1.3", "H1.4", "H1.5",
                                  "H1X","H3K9me3_SRR12349402", "K27me3",
                                  "K9me3_AJV7", "K9me3_AJV84X")

```

## 2. Construct a dataframe for each family

In order to work more efficiently, I will create one dataframe per family category.

```{r Family dataframes}

#Extract the taxonomy columns
df_tax <- dataframe_complete[, c(9, 10, 11)] #4,529,585 entries

#Get the unique class categories
classes <- sort(unique(df_tax$class))

#Get the unique family categories
families <- sort(unique(df_tax$family))

#Create an empty dataframe to store the count results
family_summary <- data.frame(family = character(), 
                             n_repeats = numeric(), 
                             stringsAsFactors = FALSE)

#Create one dataframe per family
for (category in families) {
  #Subset the dataframe for the current category
  df_category <- subset(df_tax, family == category)
  assign(paste0("df_", category), df_category)
  #Get the number of rows in the dataframe
  row_count <- nrow(df_category)
  # Add the category and row count to the result dataframe
  family_summary <- rbind(family_summary, data.frame(family = category,
                                                     n_repeats = row_count,
                                                     stringsAsFactors = FALSE))
}

#Remove the temporal "df_category" variable
rm(df_category)

```

## 3. Get the repeat counts for each family

Create one dataframe per family to store the counts of each group.

```{r Repeat counts}

# Loop through the dataframes
for (i in 1:length(families)) {
  #Create a variable that temporarily stores each dataframe
  temp_df <- get(paste0("df_", families[i]))
  #Create an empty dataframe to store the results
  temp_result_df <- data.frame(group = character(), 
                               count = integer(), 
                               stringsAsFactors = FALSE)
  #Calculate the counts for each group in the dataframe
  counts <- aggregate(temp_df$group, 
                      by = list(class = temp_df$class, 
                                family = temp_df$family, 
                                group = temp_df$group), 
                      FUN = length)
  #Rename the columns of the counts dataframe
  colnames(counts) <- c("class", "family", "group", "count")
  #Add the counts to the result dataframe
  temp_result_df <- rbind(temp_result_df, counts)
  #Order the groups from the least appearences to the most
  temp_result_df <- temp_result_df[order(temp_result_df$count), ]
  #Change rownames
  rownames(temp_result_df) <- seq(1, nrow(temp_result_df))
  #Levels and factors
  levels(temp_result_df$group) <- temp_result_df$group
  #Change the name of the dataframe
  assign(paste0("df_count_", families[i]), temp_result_df)
}

#Remove the temporal "df_category" variable
rm(temp_df)
rm(temp_result_df)

```

## 4. Plot the barplots

Create one dataframe per family to store the counts of each group.

```{r Density plots}

library(ggplot2)

#ALTERNATIVE WITHOUT GGPLOT2
#ceiling(number / 1000) * 1000
#barplot(df_count_Satellite$count, 
#             col = rgb(0.8,0.1,0.1,0.6),
#             space = 0,
#             ylim = c(0, ceiling(max(df_count_Satellite$count)/1000)*1000),
#             main = "Satellite")
#png(filename = "barplot.png", width = 600, height = 150)
#barplot(df_count_Satellite$count, 
#             col = rgb(0.8,0.1,0.1,0.6),
#             space = 0,
#             ylim = c(0, ceiling(max(df_count_Satellite$count)/1000)*1000),
#             main = "Satellite")
#graphics.off()

#Define colors for each class
class_colors <- c("#FB83DD", "#4DB8FF", "#C8C0A8", "#9898FC", "#C8F63F",
                  "#6C99DA", "#FF4444", "#5EF49E", "#FFD700", "#FE934B")
#class_colors[3] <- "#FFF700"
#class_colors[5] <- "#00A642"
#class_colors[7] <- "#008CFF"
#class_colors[8] <- "#0400FF"
#class_colors[9] <- "#9900FF"

#Set the annotation colors
annotation_class_colors <- list(class = classes, class_color = setNames(class_colors, levels(classes)))

#Color mapping
color_mapping <- setNames(annotation_class_colors$class_color, annotation_class_colors$class)

#Get a list of objects in the current environment
objects <- ls()

#Filter objects that start with "df_count"
df_names <- objects[grep("^df_count", objects)]

#Loop through the data frames and perform the desired operations
for (df_name in df_names) {
  #Get the data frame by its name
  df <- get(df_name)
  #Needed to avoid reordering of groups
  df$group <- factor(df$group, levels = df$group)
  #Create the barplot using ggplot2
  b <- ggplot(df, aes(x = group, y = count)) +
              geom_col(aes(fill = class), width = 1) +
              scale_fill_manual(values = color_mapping) +
              theme_void() +
              theme(panel.grid = element_blank(),
                    axis.text = element_blank(),
                    axis.title = element_blank(),
                    axis.ticks = element_blank()) +
              guides(fill = "none")
  #Save the plot using ggsave
  plot_name <- gsub("\\?", "_", df_name) #"?" causes ERROR
  ggsave(paste0(plot_name, "_barplot_2.png"), b, dpi = 300, width = 8, height = 2)
}

```

## 5. Plot the pie chart at class level

First, get the count and percentage data.

```{r Process class data}

#Calculate the frequencies of each class
class_counts <- table(dataframe_complete$class)

#Create a data frame for the pie chart
chart_data <- data.frame(class = names(class_counts), count = class_counts)
chart_data <- chart_data[, -2]
colnames(chart_data) <- c("class", "count")

#Calculate percentages
chart_data$percentage <- round(100 * chart_data$count / sum(chart_data$count), digits = 2)

###############################################################
#Calculate the frequencies of each family
family_counts <- table(dataframe_complete$family)

#Create a data frame for the pie chart
family_chart_data <- data.frame(family = names(family_counts), 
                                count = family_counts)
family_chart_data <- family_chart_data[, -2]
colnames(family_chart_data) <- c("family", "count")

#Calculate percentages
family_chart_data$percentage <- round(100 * family_chart_data$count / sum(family_chart_data$count), digits = 2)

###############################################################
#New colors for the pie chart
#NEW_class_colors <- c("#f36460d8", "#f4b757d8", "#ecf056d8", "#83ef5ed8", 
#                      "#5ec282d8", "#5bededd8", "#5aabf4d8", "#605af5d8",
#                      "#a75ff1d8", "#f464bcd8")

NEW_class_colors <- c("#5ef49e", "#4db8ff", "#9898fc", "#fb83dd", 
                      "#ff4444", "#fe934b", "#6c99da", "#c8f63f",
                      "#c8c0a8", "#ffd700")

#NEW_class_colors <- c("#fcdaf4", "#c6e8ff", "#e7e3d8", "#dedefd", 
#                      "#e5f6b2", "#bed1f0", "#ffd8d8", "#bdfad8",
#                      "#fdf1ac", "#faddc9")
```

Then, plot the pie chart.

```{r Pie chart class}

#ALTERNATIVE
#Calculate the frequencies of each class
#class_counts <- table(dataframe_complete$class)

#Create the pie chart
#pie(class_counts, main = "Class Distribution", col = class_colors, 
#     labels = paste0(names(class_counts), "\n", 
#                   round(100 * class_counts / sum(class_counts), 1), "%"), 
#     cex.labels = 0.8)

#Add a legend
#legend("topright", legend = names(class_counts), fill = class_colors, cex = 1,
#       inset = 0.02, text.width = 0.5)

library(ggplot2)
library(ggrepel)

chart_data$class <- factor(chart_data$class, levels = chart_data$class[order(-chart_data$percentage)])

#Create the pie chart with label placement
chart <- ggplot(chart_data, aes(x = "", y = count, fill = class)) +
         geom_bar(width = 1, stat = "identity", color = "black") +
         coord_polar("y", start = 0) +
         geom_label_repel(aes(label = paste0(class, "\n", percentage, "%")), 
                              position = position_stack(vjust = 0.5), max.overlaps = 15) +
         labs(title = "Repeats distribution in the T2T genome assembly based on the repeat class") +
         scale_fill_manual(values = NEW_class_colors) +
         theme_void()

#Display the pie chart
print(chart)

#Save the pie chart
ggsave("Albert_class_piechart.png", chart, dpi = 300, width = 8, height = 8)

```

## 6. Plot the pie chart at family level

First, get the count and percentage data. Then, plot the pie chart.

```{r Family pie chart}

library(ggplot2)
library(ggrepel)

for (i in 1:length(classes)) {
  #Subset the dataframe based on the class
  subset_data <- subset(dataframe_complete, class == classes[i])

  #Calculate the frequencies of each family
  family_counts <- table(subset_data$family)

  #Create a dataframe for the pie chart
  fam_chart_data <- data.frame(class = names(family_counts), count = family_counts)
  fam_chart_data <- fam_chart_data[, -2]
  colnames(fam_chart_data) <- c("family", "count")

  #Calculate percentages
  fam_chart_data$percentage <- round(100 * fam_chart_data$count / sum(fam_chart_data$count), digits = 2)

  #New colors for the pie chart
  n_colors <- nrow(fam_chart_data)
  fam_colors <- colorRampPalette(c(class_colors[i], "white"))(n_colors)

  #Title of the plot
  current_title <- paste0("Repeats distribution in the class ", classes[i])
  
  #Create the pie chart with label placement
  subchart <- ggplot(fam_chart_data, aes(x = "", y = count, fill = family)) +
              geom_bar(width = 1, stat = "identity", color = "black") +
              coord_polar("y", start = 0) +
              geom_label_repel(data = subset(fam_chart_data, percentage > 5),
                               aes(label = paste0(family, "\n", percentage, "%")), 
                               position = position_stack(vjust = 0.5),
                               size = 5) +
              #labs(title = current_title) +
              scale_fill_manual(values = fam_colors) +
              theme_void() +
              theme(text = element_text(size = 12)) + 
              theme(legend.position = "none")

  #Display the pie chart
  print(subchart)

  #Filename
  filename <- paste0(classes[i], "_piechart_4.png")
  
  #Save the pie chart
  ggsave(filename, subchart, dpi = 300, width = 6, height = 6)
}
```
