---
title: "KaryoploteR"
author: "Nel Marín"
date: "19/7/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Getting started

```{r Start, echo=FALSE}

#Set the working directory
setwd("~/PracticasIBMB/16_karyoploteR")

#Install the package karyoploteR
#BiocManager::install("karyoploteR")

```

## 2. Plot an ideogram

The first step when creating a karyoplot is to create the empty ideogram plot where data will later be added.

The `plotKaryotype` function does just that and returns the *karyoplot* object. We will need this object later on when we need to add data to the plot. By default it will create a karyoplot of the human genome hg19.

```{r Ideogram}

library(karyoploteR)
library(BSgenome)
library(ggplot2)

######## Plot the hg19 karyotype
hg19_kp <- plotKaryotype()

######## Plot the T2T karyotype
# Prepare the genome file (bed)
bed_genome <- as.data.frame(read.table("T2TCHM13v2.0_chrom.sizes", header = FALSE, sep="\t", stringsAsFactors=FALSE, quote=""))
bed_genome$V3 <- cbind(rep(1, times = nrow(bed_genome)))
bed_genome <- bed_genome[, c("V1", "V3", "V2")]
colnames(bed_genome) <- c("chr", "start", "end")
hs1_genome <- toGRanges(bed_genome)

# Prepare the cytobands file (bed)
bed_cytobands <- as.data.frame(read.table("T2TCHM13v2.0_Cytobands_Ideogram_UCSC.bed", header = FALSE, sep="\t", stringsAsFactors=FALSE, quote=""))
colnames(bed_cytobands) <- c("chr", "start", "end", "name", "gieStain")
hs1_cytobands <- toGRanges(bed_cytobands)

# Plot both the T2T genome and the cytobands
png("karyoplot.png", width = 1600, height = 1000)
hs1_kp <- plotKaryotype(genome = hs1_genome, cytobands = hs1_cytobands, plot.type = 6)
dev.off()

#Save as an image
png("plot.png", width = 1100, height = 700)
hs1_kp <- plotKaryotype(genome = hs1_genome, cytobands = hs1_cytobands, chromosomes= c("chr8", "chr13"))
kpAddBaseNumbers(hs1_kp, tick.dist = 10000000, tick.len = 10, tick.col="red", cex=1, minor.tick.dist = 1000000, minor.tick.len = 5, minor.tick.col = "gray")
dev.off()

png("plot1.png", width = 1800, height = 700)
hs1_kp <- plotKaryotype(genome = hs1_genome, cytobands = hs1_cytobands, chromosomes= "chr1")
kpAddBaseNumbers(hs1_kp, tick.dist = 10000000, tick.len = 10, tick.col="red", cex=1, minor.tick.dist = 1000000, minor.tick.len = 5, minor.tick.col = "gray")
dev.off()

#Save as an image
#ggsave("karyotype_chr1.png", plot = hs1_kp, device = "png")


########
getCytobandColors()
#gneg=white, gpos100=black, stalk=blue, acen=red, gvar=grey/red
########

#.libPaths()


```

## 3. Ideogram customization

In some cases it might be interesting to add a guide of the base numbering or the cytoband names to the ideogram.

```{r Customization}

# Add the base numbers to a genome
hs1_kp <- plotKaryotype(genome = hs1_genome, cytobands = hs1_cytobands) + kpAddBaseNumbers(hs1_kp)

# Add the cytoband names to an ideogram
hs1_kp <- plotKaryotype(genome = hs1_genome, cytobands = hs1_cytobands, chromosomes = "chr13") +
kpAddCytobandLabels(hs1_kp)

```

## 4. Plot continuous data

```{r Continuous}

H1.0_13 <- as.data.frame(read.table("ChIP_H1_0_NoDox_AJV32L_T47D6c2_T2T.wig_13", header = TRUE, stringsAsFactors=FALSE, quote=""))

hs1_kp <- plotKaryotype(genome = hs1_genome, cytobands = hs1_cytobands, chromosomes = "chr13")
kpArea(hs1_kp, chr="chr13", x=1:113566686, y=H1.0_13$fixedStep.chrom.chr13.start.1.step.50.span.50)
kpAxis(kp, ymin = 0, ymax=1)

####### ERROR #######

```

## 5. Alternative

```{r Alternative RIdeogram}

#install.packages("RIdeogram")
library(RIdeogram)

data(human_karyotype, package = "RIdeogram")
ideogram(human_karyotype, overlaid = bed_cytobands)
convertSVG("chromosome.svg", device = "png")
```
